Results section

### 2022-10-26 code/Result_section.Rmd
Updated results manuscript

```{r loading libraries}
library(readxl)
library(tidyverse)
library(broom)
library(ggpubr)
#library(AICcmodavg)
library(visdat)
library(rstatix)
library(gt)
library(gtsummary)
library(patchwork)
library(jtools)
```

@@

## Preface

## Introduction

This thesis is about...

## Background

## Methods and Materials

Prior to results are presented a short description of the studies conducted needs to be made.

Experiment no. 4 was the goal, but before we could test our overall hypothesis we needed to find out which combination of antibiotics and feeding scheme would results in the most NEC-like phenotype.

These research questions were investigated in 3 pre-experiments (PreExp 1-3), each with their own research questions.

@@

```{r echo=FALSE, fig.cap='Table:}
normality_result_1
```

>>>>

```{r loading libraries, message=FALSE}
library(readxl)
library(tidyverse)
library(broom)
library(ggpubr)
#library(AICcmodavg)
library(visdat)
library(rstatix)
library(gt)
library(gtsummary)
library(patchwork)
library(jtools)
```

@@

@@

```{r echo=FALSE, fig.cap='Table XX'}
normality_result_1
```

### 2022-10-26 submission/manuscript.Rmd
Updated results manuscript

```
output: html_document

```

@@

#### Rationale:

*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

## Results (decide on which voice to use, active or passive. Read other papers.)

Experiment no. 4 was the goal, but before we could test our overall hypothesis we needed to find out which combination of antibiotics and feeding scheme would results in the most NEC-like phenotype.
These research questions were investigated in 3 pre-experiments (PreExp 1-3), each with specific research questions.

#### Model feasibility
- Our first experiement (PreExp 1) was designed to test wheteher it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother. This required feeding wit a venous catheter inserted orally into the stomach. 
- `r nrow(preexp_1)` mice were born to 6 different mothers. Of these 32 survived and were included in the analysis. After fine-tuning the feeding method we we able to successfully feed all surviving (32) mice every 3 hours for 2 days (48 hours).
(Conclusion) - Visual inspection of intestines during sampling concluded that we had not created a NEC phenotype in any treatment group.
(BUT) - We found that AB-FF pups consistently looked different that CON-BF in the level of cytokine expression in the ileum, with statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES. 
- A multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val$IL-6.

#### Effects of various types of antibiotics - (Microbiome dysbiosis)
Based on the results from Preexp 1 we knew that we could perform the experimental setup without too many casualties. 
- We had seen that formula-feeding did affect the expression of cytokines, but visual inspection of the intestines showed that we had not managed to create a NEC-like phenotype. 
- After considering the results we realized that our assumptions were based on a dysbiotics microbiome in the mother leading to less immunizing milk. However, we had likely not created a dysbiosis as much as an overall depletion by administering a broad-spectered antibiotic. 
We therefore hypothezised that targeting specific strains of bacteria would lead to a more dysbiotic microbiome and allow for greater disbiosis in the offspring.

@@

#### Interaction effects between antibiotics treatment and formula feeding
- In round 3 we repeated the experimental setup with regards to MAT, adding either Vancomycin, Gentamicin or sterile water to the drinking water of pregnant "females" (use same word throughout(mothers)). We repeated the group-setup from PreExp  1 and  separated pups into breastfeeding or formula feeding, for a total of 6 groups (see overview table). 
- All pups survived the experiment and were included in the final analysis. 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

@@

### Round 4
- We treated XX mothers with Vancomycin (1 g/l) from pregnancy-day 15 until delivery, where vancomycin-mixed drinking water was replaced with regular water.
- 2 days after birth XX of the 26 pups were separated from the mother and received XX µl FvT solution or an equal amount of SM buffer (control treatment). The remaning xx of 26 pups remained with the mother for breastfeeding thoughout the study.
- Pups separated from their mother were formula fed (FF) with a solution of XXXXX milk formulation (see materials) every 3 hours for 48 hours. (futher setup description see materials and methods).
- All 26 pups survived the 48 hour study. On the final day of the study pups fasted for XX hours before given xx µl FITC Dextran (as per materials protocol). 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

@@

## Materials and methods
Describe Pre-Exp 1-3 and knowledge gained from them.

>>>>

---
title: "Results section"
author: "Kaare"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Code

# Manuscript

@@

## Results

The main goal of this thesis project was to create a NEC mouse model by combining Maternal Antibiotics Treatment with Formula-Feeding. This hypothesis was tested in a series of experiments (Exp 1-4), each successive experiment building on the results of the former. 
The following is a description of each experiment. An abridged version of each experiment is provided, followed by a detailed  description of experimental setup and results.

### Exp 1: Model feasibility

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 1. **Date of exp** 
*Antimals and experimental setup*
- Six male and 18 female C57BL/6N mice were co housed for 24 hours in 6 cages (1 male + 3 female each).
- Two weeks after mating, pregnant females were separated evenly into maternal antibiotic treatment (MAT) and water-only controls (CON). 
- Antibiotics (Gentamicin, Vancomycin, Neomycin, and Ampicillin (all 0.5 g/L)) was mixed into the drinking water of the MAT mice while controls received water only.
- Antibiotic treatment commenced on gestational day 15 and continued until delivery. 
- Each mother was housed with her offspring in separate cages during the first 3 days after birth. 
- On the 4th day, `r nrow(preexp_1)` mice, XXX (Filter `r nrow(preexp_1)` for feed type) from antibiotics treated mothers and XXX (Filter `r nrow(preexp_1)` for feed type) were divided into four experimental groups: MAT + formula-feeding (AB-FF) MAT + breastfeeding (AB-BF), Water + formula-feeding (CON-FF), and CON + breastfeeding (CON-BF).

*Feeding* **See materials and methods**
- Formula-feeding was performed by fixating the pup between thumb and forefinger, and inserting 2 cm of a venous catheter, designed for human infants, into the mouth of the pup using pliers. 
- The PICC was connected to a 1.0 ml syringe with formula. For pups >2g, 0.08 ml formula was administered (0.06 for pups ≤ 2g). 
- Feeding was done in the same order every 3 hours for the study duration, including night hours to ensuring no overnight fasting.

At 08:00 am on the day of euthanasia FORM pups received a final feeding and BF pups were separated from the mothers (halting breastfeeding). At 11:00 (three hours after the last feeding/separation), 50 µL FITC-dextran (0.045 mg/µL) was administered orally using a PICC line and a 1 mL syringe for a final dose of 750 mg/kg. Euthanasia and sample collection commenced at noon (12 am).

*Euthanasia and sample collection*
The head of the pups is cut off, and blood dripping is collected in an Eppendorf tube. The blood samples are spun down, and serum is collected and stored at -80 for cytokine measurements. Small and large intestines will be removed and evaluated macroscopically (remember photographs) for signs of NEC (intestinal segments of bowel necrosis, intestinal haemorrhage, and perforation). In addition, 1 cm of terminal ileum will be removed and stored in 4% paraformaldehyde for histological evaluation. The remaining small intestine is preserved in RNAlater for qPCR. Cecum and colon are dissected and stored in an Eppendorf tube on dry ice. 
H
istologic injury score of NEC in neonatal mice
Initially, the bowel is investigated macroscopically. Pictures are taken of the intestines in situ after abdominal incision and abdominal dissection. The location and extent of pathological areas are estimated. The severity of intestinal lesions will be assessed histopathologically in H&E-stained ileal tissue sections using an intestinal tissue injury score (0–4), where: Grade 0: intact tissue; 1: mild damage, slight submucosal and/or lamina propria separation; 2: moderate-severe separation of the submucosa and/or lamina propria and/or oedema in the submucosal and muscular layers, partial loss of villous sloughing; 3: severe separation of the submucosa and/or lamina propria region, villous sloughing and initial villus necrosis; 4: necrosis and loss of villi structure and/or transmural necrosis. Samples with histological scores of 2 or higher will be considered positive for NEC.

Samples to be collected:
•	From parent mice:
o	Faecal matter (MAT and CON mice)
•	From offspring post euthanasia
o	1-2 cm of ileum immersion-fixed in paraformaldehyde for histological analysis
o	1-2 cm of ileum snap-frozen for mesoscale + qPCR analysis
o	Whole caecum for microbiota analysis

 Of these 32 survived and were included in the analysis. After fine-tuning the feeding method we we able to successfully feed all surviving (32) mice every 3 hours for 2 days (48 hours).
(Conclusion) - Visual inspection of intestines during sampling concluded that we had not created a NEC phenotype in any treatment group.
(BUT) - We found that AB-FF pups consistently looked different that CON-BF in the level of cytokine expression in the ileum, with statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES. 
- A multiple linear regression analysis, evaluating the effects of MAT (AB or CON) and type of feed (BF of FF) separately and interactively found statistically significant effects of formula feeding for REG_MODEL_CYTOKINES$TERM == TYPEOFFEED, with the greatest effect estimate of ESTIMATE$IL-6 , P-val$IL-6.

### Experiment 2: Effects of various types of antibiotics - (Microbiome dysbiosis)

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 2.
Based on the results from Preexp 1 we knew that we could perform the experimental setup without too many casualties. 
- We had seen that formula-feeding did affect the expression of cytokines, but visual inspection of the intestines showed that we had not managed to create a NEC-like phenotype. 
- After considering the results we realized that our assumptions were based on a dysbiotics microbiome in the mother leading to less immunizing milk. However, we had likely not created a dysbiosis as much as an overall depletion by administering a broad-spectered antibiotic. 
We therefore hypothesized that targeting specific strains of bacteria would lead to a more dysbiotic microbiome and allow for greater disbiosis in the offspring.

@@

### Experiment 3 - Interaction effects between antibiotics treatment and formula feeding

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 3.
- In round 3 we repeated the experimental setup with regards to MAT, adding either Vancomycin, Gentamicin or sterile water to the drinking water of pregnant "females" (use same word throughout(mothers)). We repeated the group-setup from PreExp  1 and  separated pups into breastfeeding or formula feeding, for a total of 6 groups (see overview table). 
- All pups survived the experiment and were included in the final analysis. 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

@@

### Experiment 4

#### Rationale:
*Chen et al.*, showed that treatment with antibiotics during pregnancy(Maternal Antibiotics Treatment, **MAT**) "resulted" in mild NEC-like symptoms in the offspring. We tested the hypothesis that MAT combined with formula feeding (**FF**) would result in a worsened phenotype - [even more NEC-like than either treatment on it's own.]{.underline}

#### Aim: 
- To test whether it was possible to keep mice alive from postnatal day 3 and 48 hours onwards, while being separated from the mother.
- Whether combining a broad spectrum antibiotic with formula-feeding will result in NEC-like symptoms such as:
- Visible signs of intestinal inflammation and / or greater levels of pro-inflammatory cytokines expressed in tissue or serum

#### Conclusion:
- Survival rate: xx of `r nrow(preexp_1)` (%)
- Visaul inspection: no clear signs of intestinal inflammation
- Cytokine expression: Statistically significant differences in the level of STATISTICAL-SIGNIF-CYTOKINE-NAMES.

#### Detailed description of Exp 4.
- We treated XX mothers with Vancomycin (1 g/l) from pregnancy-day 15 until delivery, where vancomycin-mixed drinking water was replaced with regular water.
- 2 days after birth XX of the 26 pups were separated from the mother and received XX µl FvT solution or an equal amount of SM buffer (control treatment). The remaning xx of 26 pups remained with the mother for breastfeeding thoughout the study.
- Pups separated from their mother were formula fed (FF) with a solution of XXXXX milk formulation (see materials) every 3 hours for 48 hours. (futher setup description see materials and methods).
- All 26 pups survived the 48 hour study. On the final day of the study pups fasted for XX hours before given xx µl FITC Dextran (as per materials protocol). 
- 2 hours after FITC administtion pups were euthanized by decapitation and blood collected in eppendorfs for analysis of serum FITC concentration. Intestinal samples were collected for gene expression assays, histological assessment, pro-inflammatory cytokine analysis and gut microbiome sequencing.

@@
